final-project

### 2024-11-07 final-project.Rmd
table 1

```{r}
library(dplyr)
library(stringr)
library(psych)
```

@@

# further filter and select only variables used in analysis
seer <- seer_clean %>%
  filter(
    !is.na(subtype)
  ) %>%
  select(
    age,
    sex,
    year_diagnosis,
    race,
    marriage_status,
    subtype,
    t_stage = tnm,
    n_stage,
    m_stage,
    grade,
    idc_type,
    radiation,
    chemotherapy,
    surgery,
    survival_months,
    survival_status,

@@

```{r}

>>>>

```{r}
library(dplyr)
library(stringr)
library(psych)
library(gmodels)
```

@@

# there are soveral variables that need to be categorized for analysis and/filtered
seer_cat <- seer_clean %>%
   filter(
    !is.na(subtype),
    !is.na(grade),
    n_stage != "NX"
  ) %>%
  mutate(agecat = case_when(
    age %in% c("20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years") ~ 0,
    age %in% c("50-54 years", "55-59 years", "60-64 years", "65-69 years") ~ 1,
    age %in% c("70-74 years", "75-79 years", "80-84 years", "85+ years") ~ 2,
    TRUE ~ NA_real_  # Keeps other values as NA if they don’t match any of the above groups
  ))

# further filter and select only variables used in analysis
seer <- seer_cat %>%
  filter(
    !is.na(subtype),
    !is.na(grade),
    n_stage != "NX"
  ) %>%
  select(
    age = agecat,
    sex,
    year_diagnosis,
    race,
    marriage_status,
    subtype,
    n_stage,
    m_stage,
    grade,
    radiation,
    chemotherapy,
    surgery,
    survival_months,
    survival_status,

@@

```{r}
# Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
# (unadjusted associations between covariates and exposure)

## age
CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## race
CrossTable(race, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## marriage status
CrossTable(marriage_status, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## N stage
CrossTable(n_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## M stage
CrossTable(m_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## grade
CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## radiation
CrossTable(radiation, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## chemotherapy
CrossTable(chemotherapy, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## surgery
CrossTable(surgery, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

```{r}
# Table 2: Clinicopathological Characteristics by survival
# (Unadjusted associations between covariates and survival)
